---
title: "COVID-19 Cases in Switzerland"
output: 
  github_document: default
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
theme_set(theme_bw(base_size=14))
```

# Data

Read in the data.

```{r readIn}
dat <- read_csv("COVID19_Fallzahlen_CH_total.csv")
head(dat)
```

Fill in data for missing dates. Decide how to handle incomplete data from last (current) day in case of summary stats for whole Switzerland.

- `removeLast=TRUE` removes last (current day)
- `removeLast=FALSE` transfer the values from previous day for missing cantons

```{r removeLast}
removeLast <- FALSE
```


```{r fillOffset}
maxDate <- max(dat$date)
if (removeLast) {
  offset <- 0
} else {
  offset <- 1
}
```


```{r cleanData}
dat <- dat %>% 
  group_by(abbreviation_canton_and_fl, date) %>%          # group by region and date
  summarize_if(is.numeric, max, na.rm=TRUE) %>%           # select max value (if there are several per day)
  mutate_if(is.numeric, na_if, -Inf) %>%                  # replace the -Inf with NA to enable fill
  complete(date=seq.Date(min(date), max(date), by=1)) %>% # add missing dates (values filled with NA's)
  fill(-c(date, abbreviation_canton_and_fl)) %>%          # fill NA's with previous value
  mutate_if(is.numeric, replace_na, 0)                    # replace NA's at the beginning of the time-series
```

Select data for one canton (BS).

```{r dataOneCanton}
dat %>%
  filter(abbreviation_canton_and_fl=="BS") %>%               # filter specific region
  as.data.frame()
```


## Plotting number of cases

For specific canton,  Basel Stadt (BS) as an example.

```{r plotOneCanton}
dat %>%
  filter(abbreviation_canton_and_fl=="BS") %>%               # filter specific region
  ggplot(aes(x=date, y=ncumul_conf)) +
  geom_line(colour="steelblue", size=1) + scale_y_log10() +
  xlab("") + ylab("No. of confirmed cases") + ggtitle("Canton Basel Stadt (BS)")
```

For several cantons,  Basel Land (BL) and Basel Stadt (BS) as an example.

```{r plotMultipleCantons}
dat %>%
  filter(abbreviation_canton_and_fl %in% c("BL","BS")) %>%    # filter specific region(s)
  ggplot(aes(x=date, y=ncumul_conf)) +
  geom_line(aes(colour=abbreviation_canton_and_fl), size=1) + scale_y_log10() +
  xlab("") + ylab("No. of confirmed cases") + ggtitle("Cantons Basel Land (BL) and Stadt (BS)") +
  theme(legend.title=element_blank())
```


## Whole Switzerland

```{r plotSwitzerland}
dat %>%
  complete(date=seq.Date(min(date), maxDate, by=1)) %>%   # add missing dates (values filled with NA's) 
  fill(-c(date, abbreviation_canton_and_fl)) %>%          # fill NA's with previous value
  mutate_if(is.numeric, replace_na, 0) %>%                # replace NA's at the beginning of the time-series
  group_by(date) %>%                                      # group by date and
  summarize_if(is.numeric, sum) %>%                       # sum-up all cases per date from all regions
  filter(date < maxDate+offset) %>%                       # if set, remove incomplete data for last day
  ggplot(aes(x=date, y=ncumul_conf)) +
  geom_line(colour="steelblue", size=1) + scale_y_log10() +
  xlab("") + ylab("No. of confirmed cases") + ggtitle("Switzerland")
```


# Doubling rate (simple)

## Per canton

```{r doublingPerCanton}
dat %>% 
  arrange(date) %>%                                              # order by date
  group_by(abbreviation_canton_and_fl) %>%                       # group by region
  summarize_if(is.numeric, function(x) {sum(x >= x[length(x)]/2)-1}) %>% # count number of days to double
  rename_at(vars(starts_with("ncumul")), str_replace, pattern="ncumul", "daysToDouble") %>%
  as.data.frame()
```

## Whole Switzerland

```{r doublingSwitzerland}
dat %>% 
  complete(date=seq.Date(min(date), maxDate, by=1)) %>%          # add missing dates (values filled with NA's) 
  fill(-c(date, abbreviation_canton_and_fl)) %>%                 # fill NA's with previous value
  mutate_if(is.numeric, replace_na, 0) %>%                       # replace NA's at the beginning of the time-series
  arrange(date) %>%                                              # order by date
  group_by(date) %>%                                             # group by date and
  summarize_if(is.numeric, sum) %>%                              # sum-up all cases per date from all regions
  filter(date < maxDate+offset) %>%                              # if set, remove incomplete data for last day
  summarize_if(is.numeric, function(x) {sum(x >= x[length(x)]/2) -1}) %>% # count number of days to double
  rename_at(vars(starts_with("ncumul")), str_replace, pattern="ncumul", "daysToDouble") %>%
  as.data.frame()
```

